---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r, echo=FALSE,message=FALSE}
library(dplyr)
library(ggplot2)
```

```{r}
mydata.df=read.csv("./activity.csv",header=T,
	colClasses=c("integer","Date","integer"))
SumEachDay.df=aggregate(mydata.df$steps,by=list(Day=mydata.df$date),FUN=sum)
```

##
## What is mean total number of steps taken per day?
##
Below is the table for mean & median for number of step taken each day:
```{r, echo=FALSE}
hist(SumEachDay.df$x)
```

## What is the average daily activity pattern?

Below is the table for mean & median for number of step taken each day:
```{r showtable, result="asis"}
MeanMedian.df=mydata.df %>% 
	group_by (date) %>% 
	summarise(mean=mean(steps,na.rm=T), median=median(steps,na.rm=T))
MeanMedian.df
```

Below is the plot with Ave number of steps vs 5 mins interval:
```{r}
MeanFiveMinutes=mydata.df %>% group_by (interval) %>% summarise(average=mean(steps,na.rm=T))
ggplot(data=MeanFiveMinutes) + geom_line(aes(x = interval, y = average)) + xlab('Mins') + ylab('Average')
```

```{r}
myinterval=subset(MeanFiveMinutes, average==max(MeanFiveMinutes$average),
        select=interval)
```
`r myinterval` is the max number of steps.

##
## Inputing missing values
##
```{r, echo=FALSE}
TotalRowNA=nrow(subset(mydata.df, is.na(steps)))
```
Total number of missing values in the dataset is `r TotalRowNA`.

```{r, message=FALSE,warning=FALSE}
library(reshape)
mydataCast.df=cast(mydata.df,date~interval,value="steps")

for (x in 1:nrow(mydataCast.df)){
        myvalue = as.numeric(mydataCast.df[x,-1])
        myvalue[is.na(myvalue)] = MeanFiveMinutes$average
        mydataCast.df[x,-1] <- myvalue
}
```

The NA value in dataset are replaced with 5mins interval  mean value.

```{r, echo=FALSE}
mydataClean.df= mydataCast.df %>%
                melt(id=date) %>%
                arrange(date,interval)

colnames(mydataClean.df)[2]="steps"

```


```{r}
SumEachDayClean.df=aggregate(mydataClean.df$steps, by=list(Day=mydataClean.df$date),FUN=sum)
ReportClean=mydataClean.df %>% group_by (date) %>% summarise(mean=mean(steps,na.rm=T), median=median(steps,na.rm=T))
```

##Below is the table for mean & median for each day with NA "cleaned" based on above strategy:
```{r, echo=FALSE}
hist(SumEachDayClean.df$x)
```

Below is the table for mean & median for number of step taken each day with NA filled up based on above strategy :
```{r}
ReportClean.df=mydataClean.df %>% group_by (date) %>% summarise(mean=mean(steps,na.rm=T), median=median(steps,na.rm=T))
ReportClean.df
```


```{r, echo=FALSE,message=FALSE}

library(lubridate)
mydata.df$day = wday(mydata.df$date) 

# --- conflict with dplyr and cause uncertainty to summarize function -------
#library(plyr)
#mydata.df$day = wday(mydata.df$date) %>% factor %>% revalue ( #c("1"="weekday","2"="weekday","3"="weekday","4"="weekday","5"="weekday", "6"="weekend", #"0"="weekend"))

Weekdayclass=1:5
Weekendclass=6:7

mydata.df$day[ mydata.df$day %in% Weekdayclass ] <- "weekday"
mydata.df$day[ mydata.df$day %in% Weekendclass ] <- "weekend"

MeanFiveMinutesByDay=mydata.df %>% group_by (interval,day) %>% summarise(average=mean(steps,na.rm=T))
myweekday=subset(MeanFiveMinutesByDay, day=="weekday", select=c(interval, average))
myweekend=subset(MeanFiveMinutesByDay, day=="weekend", select=c(interval, average))
ggplot() + geom_line(data=myweekday,aes(x = interval, y = average),color="red") + 
	geom_line(data=myweekend,aes(x = interval, y = average),color="black") + 
	xlab('Mins') + ylab('Average')  + theme(legend.position="bottom") + 
  theme(legend.title=element_blank()) + guides(fill=TRUE)

```

# Are there differences in activity patterns between weekdays and weekends?
# 
## From the graph, there are high mean steps between 0500 - 1000 on weekday. After 1000 onwards,  there are more steps taken over weekend.
